Estimation of Stochastic Wilson and Cowan Neural Field Model using Expectation Maximisation Algorithm

### 2011-08-18 ltx/data_paper/DataPaper.tex
edits to correlation analysis results and methods

The cross-correlation is further simplified by recognizing that the first term on right hand side of equation~\eqref{eq:EM-ExpectationToCancelNoise} can be written as 

@@

	f\left(v_t(\mathbf{r})\right) \approx \varsigma v_t(\mathbf{r})

@@

@@

To isolate the kernel, the order of the convolution and cross-correlation is reversed by recognizing that a property $(a \ast b)(\boldsymbol\iota) \star c(\boldsymbol\iota) = a(-\boldsymbol\iota)\ast(b \star c)(\boldsymbol\iota)$ \parham{(see Appendix for proof)}. Therefore,

@@

From \eqref{eq:EM-Fourier_TF_of_Kernel} it can be observed that an error in our guess for $\xi$ will result in a dilation or contraction of the estimated kernel support and an incorrect guess of the observation noise will result in a distortion in the shape of the kernel, however if the signal to noise ratio is high such a distortion becomes insignificant. The denominator in \eqref{eq:EM-Fourier_TF_of_Kernel} is in fact the power spectral density of the noise-free observations and is a non-negative and real quantity. This  is known as Weiner-Khintchine (W-K) theorem in signal processing which states that the power spectral density of the wide sense stationary process is the Fourier transform of its autocorrelation function \cite{Ricker2003}. The numerator is also known as the cross-spectrum or cross-spectral density. Since the connectivity kernel is a real function, the following relation holds \cite{Bracewell2000}

@@

From equation~\ref{eq:KernelSolution} it is clear that the an initial estimate of the synaptic time constant is required to estimate the shape of the connectivity kernel from data. \dean{Maybe we can say something like:} From equation~\ref{eq:EM-Fourier_TF_of_Kernel}, it can be seen the an error in the estimate of the value of the synaptic time constant will yield either shift up or down in the frequency domain. Since the goal of the procedure is to constrain the positioning of connectivity kernel basis functions, the initial estimate of the time constant should be conservatively \dean{higher (or lower?)}, which will result in a dilation of shape of the kernel (shift lower in the frequency domain). If the shape of the kernel estimate is wider than the actual kernel, then we can be sure that the positioning of the basis functions will capture the connectivity structure once the basis function coefficients have been estimated. If the kernel shape is too narrow, then the parameter values associated with the basis function can not compensate.

In addition to the synaptic time constant, an initial guess for the observation noise variance is also required.

In doing so, the unknown parameters $\sigma_{\epsilon}^2$ and $\xi$ are required and should be guessed beforehand. To form a more accurate initial guess of the unknown parameters and explore the effect of the chosen initial guess on the estimated supports several tests were performed. First consider \eqref{eq:EM-MMGFourier2} for estimating the disturbance covariance function. As mentioned before, the power spectral densities are non-negative. Therefore, the minimum values of $\mathcal F \{\mathbf E\left[ R_{y_ty_t}\right] \}$ and $\mathcal F \{\mathbf E\left[ R_{y_{t+1}y_{t+1}}\right]\}$ should be chosen (\dean{these aren't chosen, they are calculated from data, right? Not sure what you mean. Do mean the initial guess of `$\sigma^2_{\varepsilon}$' should be chosen such that ...? or $\sigma^2_{\varepsilon}$ is bounded by...}) as the higher limit for the initial guess of $\sigma^2_{\varepsilon}$. The minimum value of the power spectral densities was found to be 0.23.

>>>>

The cross-correlation is further simplified by recognising that the first term on right hand side of equation~\eqref{eq:EM-ExpectationToCancelNoise} can be written as 

@@

	f\left(v_t(\mathbf{r})\right) \approx \varsigma v_t(\mathbf{r}).

@@

\dean{I think we should use this activation function (this will fix the distortion across the $x=0,y=0$ axis):
The function linearised about the firing threshold is 
	\hat{f}(v_t\left(\mathbf{r}\right)) &= f(v_0) + f'(v_0)(v_t\left(\mathbf{r}\right) - v_0) \\
	&= \frac{2 + \varsigma(v_t\left(\mathbf{r}\right) - v_0)}{4}. \label{eq:LinearActivationFunction} 
\end{align}}

@@

To isolate the kernel, the order of the convolution and cross-correlation is reversed by recognising that a property $(a \ast b)(\boldsymbol\iota) \star c(\boldsymbol\iota) = a(-\boldsymbol\iota)\ast(b \star c)(\boldsymbol\iota)$ \parham{(see Appendix for proof)}. Therefore,

@@

From \eqref{eq:EM-Fourier_TF_of_Kernel} it can be observed that an error in the initial estimate of $\xi$ will result in a dilation or contraction of the estimated kernel support and an incorrect guess of the observation noise will result in a distortion in the zero frequency (offset) component of the kernel, however if the signal-to-noise ratio is high such a distortion becomes insignificant.

The denominator of equation \eqref{eq:EM-Fourier_TF_of_Kernel} is equivalent to the power spectral density (PSD) of the noise-free observations. By the Weiner-Khintchine theorem (WKT) the PSD is non-negative and a real quantity. The WKT states that the power spectral density of the wide sense stationary process is the Fourier transform of its autocorrelation function \cite{Ricker2003}. \dean{The WKT allows us establish the bounds}
\begin{align}
	\mathcal{F}\{(\mathbf{E}\left[R_{y_ty_t}(\boldsymbol\iota)\right] - \sigma_{\varepsilon}^2 \delta(\boldsymbol\iota))\}\ge&0 \\
	\implies & \mathcal{F}\{(\mathbf{E}\left[R_{y_ty_t}(\boldsymbol\iota)\right]\}\ge\mathcal{F}\{\sigma_{\varepsilon}^2 \delta(\boldsymbol\iota))\}\ge0 \\
	\implies & \min_f\mathcal{F}\{(\mathbf{E}\left[R_{y_ty_t}(\boldsymbol\iota)\right]\}\ge\sigma_{\varepsilon}^2\ge0. \label{eq:BoundOnObsVariance}
\end{align}
The numerator is also known as the cross-spectrum or cross-spectral density.

Since the connectivity kernel is a real function, the following relation holds \cite{Bracewell2000}

@@

From equation~\ref{eq:KernelSolution}, it is clear that the an initial estimate of the synaptic time constant is required to estimate the shape of the connectivity kernel from data. \dean{Maybe we can say something like:} From equation~\ref{eq:EM-Fourier_TF_of_Kernel}, it can be seen the an error in the estimate of the value of the synaptic time constant will yield either shift up or down in the frequency domain, which leads to a constriction or dial. Since the goal of the procedure is to constrain the positioning of connectivity kernel basis functions, the initial estimate of the time constant should be conservatively \dean{higher (or lower?)}, which will result in a dilation of shape of the kernel (shift lower in the frequency domain). If the shape of the kernel estimate is wider than the actual kernel, then we can be sure that the positioning of the basis functions will capture the connectivity structure once the basis function coefficients have been estimated. If the kernel shape is too narrow, then the parameter values associated with the basis function can not compensate.

In addition to the synaptic time constant, an initial guess for the observation noise variance, $\sigma_{\epsilon}^2$, is also required. From equation~\eqref{eq:BoundOnObsVariance}, the value of $\sigma_{\epsilon}^2$ is bounded. Using the artificial data, the upper bound on $\sigma_{\epsilon}^2$ (the minimum value of the power spectral densities) was found to be approximately $0.23$.

To explored the sensitivity of the initial guess of $\sigma_{\epsilon}^2$, several tests were performed. \dean{The connectivity kernel should come first when presenting the results, since the derivation is presented first in the method section.} First consider \eqref{eq:EM-MMGFourier2} for estimating the disturbance covariance function.
